% -*- root: main.tex -*-
\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Implementation}
\label{sec:implementation}


\subsection{A Brief Review of Agda}
\label{ssec:agda-definition}

\Agda is an interactive system for constructing proofs and programs,
based on Martin-L\"{o}f's type theory and extended with records, parametrised
modules, among other features.
One of the main strengths of \Agda is its support for writing proofs,
which we shall call \Agda's proof engine. It consists of: support for
inductively defined types, including inductive families, and function
definitions using pattern matching on such types, normalisation during type-
checking, commands for refining proof terms, coverage checker and termination
checker.

The inductive approach for representing classical propositional logic
is better because we benefit from \Agda's proof engine and its Unicode
support that allows us writing proofs similar as we find in math text books.

In \Agda, we define a proposition formula
as an inductive type using the keyword \texttt{data} extending the
its definition in \cite{Altenkirch2015}.  
The \name{PropFormula} data type stands for propositions
as we define in section \ref{sec:preliminaries}.
We include every connective and logic constant as a constructor.
We also define the constructor for atomic propositions denoted by
\verb!Var!. This atomic expressions are represented using a natural 
number $i$ bounded by a size of the universe $n$ for those atomic 
propositions (in \Agda, we use the data type \verb!Fin! from the 
standard  library for such a purpose).

\begin{verbatim}
data PropFormula : Set where
  Var : Fin n → PropFormula
  ⊤ : PropFormula
  ⊥ : PropFormula
  _∧_ _∨_ _⇒_ _⇔_ : (φ ψ : PropFormula) → PropFormula
  ¬_ : (φ : PropFormula) → PropFormula
\end{verbatim}

\begin{remark}
The underscores in the above definition denote holes for a function of arity equivalent to the number of such holes.
\end{remark}

\begin{remark}
The type notation
\begin{verbatim}
_⇔_ : (φ ψ : PropFormula) → PropFormula
\end{verbatim}
is a syntax sugar for the type
\begin{verbatim}
PropFormula → PropFormula → PropFormula.
\end{verbatim}
\end{remark}

\begin{example}
Using Unicode translating a proposition from a textbook to \Agda 
using the above data type is straightforward. 
\begin{verbatim}
r : PropFormula
p = Var (# 0)
q = Var (# 1)
r = Var (# 2)

φ : PropFormula
φ = (¬ r ∨ p) ∧ (¬ r ∨ q)
\end{verbatim}
\end{example}

\end{document}